
#Numerical Variables Graphs 

Loading dataset
```{r}
co2emissions=read.csv("CO2 Emissions_Canada.csv",header=FALSE)
```

Plotting a boxplot for Engine Size
```{r}
co2emissions$V12<-as.numeric(co2emissions$V12) #converting to numeric since first object is column name in string format
co2emissions$V4<-as.numeric(co2emissions$V4)
#co2emissions

min(co2emissions$V4,na.rm="TRUE")
max(co2emissions$V4,na.rm="TRUE")
boxplot(co2emissions$V4,ylim=c(0.5,9),main="Engine Size",horizontal = TRUE,border="brown",xlab="Engine size in Litres",col="bisque",na.rm=TRUE)
```
Plotting a boxplot for Cylinders

```{r}
co2emissions$V5<-as.numeric(co2emissions$V5)
boxplot(co2emissions$V5,main="Cylinders",horizontal=TRUE,xlab="Numer of Cylinders",border="deepskyblue4",col="darkslategray1",na.rm=TRUE)
```

Plotting a boxplot for Fuel Consumption in City Roads
```{r}
co2emissions$V8<-as.numeric(co2emissions$V8)
boxplot(co2emissions$V8,main="Fuel Consumption in City Roads",horizontal=TRUE,xlab="Fuel consumption (Litres/100km)",border="chocolate4",col="lightpink",na.rm=TRUE)
```
Plotting a boxplot for Fuel Consumption in Highways
```{r}
co2emissions$V9<-as.numeric(co2emissions$V9)
boxplot(co2emissions$V9,main="Fuel Consumption in Highways",horizontal=TRUE,xlab="Fuel consumption (Litres/100km)",border="lightgoldenrod4",col="lightgoldenrod1",na.rm=TRUE)
```
Plotting a boxplot for Combined Fuel Consumption (City Roads and Highways)
```{r}
co2emissions$V10<-as.numeric(co2emissions$V10)
boxplot(co2emissions$V10,main="Combined Fuel Consumption",horizontal=TRUE,xlab="Fuel consumption (Litres/100km)",border="orangered4",col="orange3",na.rm=TRUE)
```

#Categorical Variables Graphs 
```{r}
data = read.csv("CO2 Emissions_Canada.csv")

summary_data <- tapply(data$CO2.Emissions.g.km., data$Cylinders,
                       FUN = mean, na.rm = TRUE)
summary_data

library(RColorBrewer)
coul <- brewer.pal(5, "Set2") 

barplot(summary_data, xlab = "Cylinders",
        main = "CO2 emission",
        col = coul,
        beside = TRUE,
        legend.text = rownames(summary_data),
        args.legend = list(title = "Cylinders", x = "topright",
                           inset = c(-0.20, 0)))


   
```

```{r}


summary_data <- tapply(data$CO2.Emissions.g.km., data$Transmission,
                       FUN = mean, na.rm = TRUE)
summary_data

library(RColorBrewer)
coul <- brewer.pal(5, "Set2") 

barplot(summary_data, xlab = "Transmission Type",
        main = "CO2 emission",
        col = coul,
        beside = TRUE,
        args.legend = list(title = "Transmission Type", x = "topright"), 
        las=2)


```

```{r}
data2 = read.csv("CO2_Emissions_Canada_1.csv")

summary_data1 <- tapply(data2$CO2.Emissions.g.km., data2$Vehicle.Class,
                       FUN = mean, na.rm = TRUE)
summary_data1

par(mar=c(3, 16, 3, 1)) 

barplot(summary_data1,
        main = "CO2 emission per Vehicle class",
        col = coul,
        las=2,
        horiz=TRUE)


```

```{r}


summary_data <- tapply(data$CO2.Emissions.g.km., data$Fuel.Type,
                       FUN = mean, na.rm = TRUE)
summary_data



barplot(summary_data, xlab = "Fuel Type",
        main = "CO2 emission",
        col = coul,
        beside = TRUE,
        legend.text = rownames(summary_data))


```



```{r}
set.seed(4555)

library(caret)
library(fastDummies)
library(corrplot)
library(dplyr)
raw_data <- read.csv("CO2 Emissions_Canada.csv")

raw_data$Vehicle.Class[raw_data$Vehicle.Class == "SUV - SMALL"] <- "SUV"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "SUV - STANDARD"] <- "SUV"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "PICKUP TRUCK - STANDARD"] <- "PICKUP TRUCK"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "PICKUP TRUCK - SMALL"] <- "PICKUP TRUCK"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "VAN - CARGO"] <- "VAN"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "VAN - PASSENGER"] <- "VAN"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "MINIVAN"] <- "VAN"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "STATION WAGON - MID-SIZE"] <- "STATION WAGON"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "STATION WAGON - SMALL"] <- "STATION WAGON"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "MINICOMPACT"] <- "COMPACT"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "SUBCOMPACT"] <- "COMPACT"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "MID-SIZE"] <- "SEDAN"
raw_data$Vehicle.Class[raw_data$Vehicle.Class == "FULL-SIZE"] <- "SEDAN"


dummy_data <- dummy_cols(raw_data, select_columns = c("Make","Vehicle.Class","Transmission","Fuel.Type"))
dummy_data <- dummy_data[,-c(1,2,3,5,6,7)]
dummy_data <- dummy_data[,-c(2,3,5)]

dummy_data.cor <- cor(dummy_data)

trainIndex <- createDataPartition(dummy_data$CO2.Emissions.g.km., p = 0.8, list = FALSE, times = 1)

train_data <- dummy_data[trainIndex,]
test_data <- dummy_data[-trainIndex,]

trainx <- select(train_data,-CO2.Emissions.g.km.)
trainx <- as.matrix(trainx)
trainy <- train_data$CO2.Emissions.g.km.


test_response <- test_data$CO2.Emissions.g.km.
testx <- select(test_data,-CO2.Emissions.g.km.)
testx <- as.matrix(testx)

```

```{r}

### Source: https://towardsdatascience.com/how-to-create-a-correlation-matrix-with-too-many-variables-309cc0c0a57
### Courtesy of Catherine Williams
corr_simple <- function(data=df,sig=0.5){
  #convert data to numeric in order to run correlations
  #convert to factor first to keep the integrity of the data - each value will become a number rather than turn into NA
  df_cor <- data %>% mutate_if(is.character, as.factor)
  df_cor <- df_cor %>% mutate_if(is.factor, as.numeric)  #run a correlation and drop the insignificant ones
  corr <- cor(df_cor)
  #prepare to drop duplicates and correlations of 1     
  corr[lower.tri(corr,diag=TRUE)] <- NA 
  #drop perfect correlations
  corr[corr == 1] <- NA   #turn into a 3-column table
  corr <- as.data.frame(as.table(corr))
  #remove the NA values from above 
  corr <- na.omit(corr)   #select significant values  
  corr <- subset(corr, abs(Freq) > sig) 
  #sort by highest correlation
  corr <- corr[order(-abs(corr$Freq)),]   #print table
  print(corr)  #turn corr back into matrix in order to plot with corrplot
  mtx_corr <- reshape2::acast(corr, Var1~Var2, value.var="Freq")
  
  #plot correlations visually
  corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
}

corr_simple(data = dummy_data, sig = 0.4)

```

### Linear Model

```{r}

linear_model <- lm(CO2.Emissions.g.km. ~ ., data = train_data)
summary(linear_model)

lin_pred <- predict(linear_model,new_data = test_data)
lin_test_mse <- mean((test_response-lin_pred)^2)
lin_test_mse

AIC(linear_model)
BIC(linear_model)

library(caret)
data.frame(
  R2 = R2(lin_pred, test_response),
  RMSE = RMSE(lin_pred, test_response),
  MAE = MAE(lin_pred, test_response))
```
### Lasso
```{r}
library(glmnet)

lasso_cross <- cv.glmnet(trainx,trainy, alpha = 1)
plot(lasso_cross)
best_lambda <- lasso_cross$lambda.min
lasso_model <- lasso_cross$glmnet.fit
lasso_coef <- coef.glmnet(lasso_model, s = best_lambda)
lasso_coef

lasso_pred <- predict(lasso_model, s=best_lambda, newx = testx)

lasso_mse <- mean((test_response-lasso_pred)^2)
lasso_mse

library(caret)
data.frame(
  R2 = R2(lasso_pred, test_response),
  RMSE = RMSE(lasso_pred, test_response),
  MAE = MAE(lasso_pred, test_response)
)
```
### Ridge
```{r}

ridge_cross <- cv.glmnet(trainx,trainy, alpha = 0)
plot(ridge_cross)
best_lambda <- ridge_cross$lambda.min
ridge_model <- ridge_cross$glmnet.fit
ridge_coef <- coef.glmnet(ridge_model, s = best_lambda)
ridge_coef

ridge_pred <- predict(ridge_model, s=best_lambda, newx = testx)

ridge_mse <- mean((test_response-ridge_pred)^2)
ridge_mse

library(caret)
data.frame(
  R2 = R2(ridge_pred, test_response),
  RMSE = RMSE(ridge_pred, test_response),
  MAE = MAE(ridge_pred, test_response)
)
```

```{r}
library(randomForest)

rforest_model <- randomForest(x = trainx, y = trainy)
rforest_pred <- predict(rforest_model, testx)
rforest_mse <- mean((test_response-rforest_pred)^2)
rforest_mse

library(caret)
data.frame(
  R2 = R2(rforest_pred, test_response),
  RMSE = RMSE(rforest_pred, test_response),
  MAE = MAE(rforest_pred, test_response)
)
```

#Gradient boosting 

```{r}
library(gbm)

boost_model=gbm(CO2.Emissions.g.km. ~ ., data = train_data,distribution = "gaussian",n.trees = 10000,
                  shrinkage = 0.01, interaction.depth = 4)
boost_pred <- predict(boost_model, test_data)
boost_mse <- mean((test_response-boost_pred)^2)
boost_mse

library(caret)
data.frame(
  R2 = R2(boost_pred, test_response),
  RMSE = RMSE(boost_pred, test_response),
  MAE = MAE(boost_pred, test_response)
)
```


























